The Symmetry of Partner Modelling

### 2016-02-04 mutual-modeling.tex
Text reflow at 80 characters

\title{The symmetry of partner modelling}

@@

\begin{abstract}
    Collaborative learning has often been associated to the construction of  a shared   understanding of the situation at hand. The psycholinguistics mechanisms
    at work while establishing common grounds are the object of scientific controversy. We postulate that collaborative tasks require some level of mutual 
    modeling, i.e. that each partner needs some model of what the other partners know/want/intend at a given time.  We use the term 'some model' to stress the fact that this model is not necessarily details or complete, but that we acquire some representations of the persons we interact with. The question we address is: 
Does the quality of the partner model depend upon the modeler's ability to represent his or her partner, upon the modelee's ability to make his state clear to the modeler or upon the quality of their interactions? We address this question by analyzing differences between the  accuracy of the models built by different team members. We therefore report 5 experiments on collaborative problem solving or collaborative learning that vary in terms of tasks (how important it is to build an accurate model) and settings (how difficult it is to build an accurate model). In 4 studies, the accuracy of the model that A built about B was correlated with the accuracy of the model that B built about A,  which seems to imply that the quality of interactions matters more than individual abilities when building mutual models.  However, these finding do not discard the fact individual abilities also  participate in the quality of modellng process .

@@

\vspace{2em}
In summary, the controversy around the cognitive depth of shared understanding pertains to psycholinguistic, which investigates 'natural conversations, but  in collaborative learning, the situation is different; tasks are precisely designed for requirint a deeper negotiation fo meaning.  The craft of CSCL is to design collaborative situations that create interdependence, avoid groupthink, and allows students to detect any 'illusion of shared understanding' (Cherubini et al, 2005). Our question is hence not  anymore"do peers build a shared understanding ?", but becomes "when peers have to build a shared understanding, how do they do it ?".  This question addresses the mechanisms of grounding, in which the basic sequence is: make a proposition, detect misunderstanding and, if any, repair them.  This paper focuses on the middle part, the detection of misunderstandings.  Detecting a misunderstanding means that the emitter of a message identifies a mismatch between his communicative intention and the way his message is understood by his or her partner. Detecting one peer's understanding is investigate hereafter under the general umbrella of partner modelling.

Cherubini, M., Van Der Pol, J., & Dillenbourg, P. (2005). Grounding is not shared understanding: Distinguishing grounding at an utterance and knowledge level. In CONTEXT'05, the Fifth International and Interdisciplinary Conference on Modeling and Using Context (pp. 11-23).

@@

We have however to bring some nuances to the social viewpoint ($\mathcal{H}_{3}).  The main feature that can be associated to the team level in this experiment is probably not the quality of their verbal interactions per se (they interact mostly by drawing on a shared map), but rather the consistency of  the spatial exploration strategy: a clear strategy facilitates memorizing one's partner path.  One could argue whether the team strategy can be dissociated from the team interactions quality or constitues one of its components.

@@

Like the previous one, this study leads us to refine what we mean by quality of interactions in $\mathcal{H}_{3}$. We  expected that the video channel would help peers building a more accurate model of the each other's emotions.  The results show the opposite: peers in the audio-only condition built more accurate models, probably because they concentrated more on the shared concept map. This confirms other studies that revealed that viewing what one's partner sees (shared graphical editor) is more important
than seeing each other~\citep{gaver1993one,anderson1997impact}. Therefore, what is meant by the quality of interactions is $\mathcal{H}_{3} is more than the linguistic features of dialogue but includes the way these interactions are articulated to the task.

@@

We found symmetry of the mutual model on 4 studies out of 5. In study 4, this only applies to the control group (having read the same text before), which corresponds to the situation of the 3 first studies.  Finding this symmetry per se constitutes an interesting result as we don't know earlier studies who have established this relationship. The problem is that this symmetry does not allow use to discriminate  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The second hypothesis, the skill of helping one's peer to model oneself, is questioned by study 3 where C could be modelled  accurately by A and not accurately by B. The same hypothesis is however supported by the results of study 2, and indirectly by study 4.  This does not discard \mathcal{H}_{1}$: even when we found significant correlations, they were all below 0.50, not around 0.90.  Hence, even of the quality of social interaction matters, there is obviously a large part of individual variance within teams.

>>>>

\title{The Symmetry of Partner Modelling}

@@

\begin{abstract}

Collaborative learning has often been associated to the construction of  a
shared   understanding of the situation at hand. The psycholinguistics
mechanisms at work while establishing common grounds are the object of
scientific controversy. We postulate that collaborative tasks require some
level of mutual modeling, i.e. that each partner needs some model of what
the other partners know/want/intend at a given time.  We use the term 'some
model' to stress the fact that this model is not necessarily details or
complete, but that we acquire some representations of the persons we
interact with. The question we address is: Does the quality of the partner
model depend upon the modeler's ability to represent his or her partner,
upon the modelee's ability to make his state clear to the modeler or upon
the quality of their interactions? We address this question by analyzing
differences between the  accuracy of the models built by different team
members. We therefore report 5 experiments on collaborative problem solving
or collaborative learning that vary in terms of tasks (how important it is
to build an accurate model) and settings (how difficult it is to build an
accurate model). In 4 studies, the accuracy of the model that A built about
B was correlated with the accuracy of the model that B built about A,  which
seems to imply that the quality of interactions matters more than individual
abilities when building mutual models.  However, these finding do not
discard the fact individual abilities also  participate in the quality of
modellng process .

@@

\vspace{2em}

In summary, the controversy around the cognitive depth of shared understanding
pertains to psycholinguistic, which investigates 'natural conversations, but  in
collaborative learning, the situation is different; tasks are precisely designed
for requirint a deeper negotiation fo meaning.  The craft of CSCL is to design
collaborative situations that create interdependence, avoid groupthink, and
allows students to detect any 'illusion of shared understanding' (Cherubini et
al, 2005). Our question is hence not  anymore"do peers build a shared
understanding ?", but becomes "when peers have to build a shared understanding,
how do they do it ?".  This question addresses the mechanisms of grounding, in
which the basic sequence is: make a proposition, detect misunderstanding and, if
any, repair them.  This paper focuses on the middle part, the detection of
misunderstandings.  Detecting a misunderstanding means that the emitter of a
message identifies a mismatch between his communicative intention and the way
his message is understood by his or her partner. Detecting one peer's
understanding is investigate hereafter under the general umbrella of partner
modelling.

%TODO
Cherubini, M., Van Der Pol, J., \& Dillenbourg, P. (2005). Grounding is not
shared understanding: Distinguishing grounding at an utterance and knowledge
level. In CONTEXT'05, the Fifth International and Interdisciplinary Conference
on Modeling and Using Context (pp. 11-23).

@@

We have however to bring some nuances to the social viewpoint
($\mathcal{H}_{3}$).  The main feature that can be associated to the team level
in this experiment is probably not the quality of their verbal interactions per
se (they interact mostly by drawing on a shared map), but rather the consistency
of  the spatial exploration strategy: a clear strategy facilitates memorizing
one's partner path.  One could argue whether the team strategy can be
dissociated from the team interactions quality or constitues one of its
components.

@@

Like the previous one, this study leads us to refine what we mean by quality of
interactions in $\mathcal{H}_{3}$. We  expected that the video channel would
help peers building a more accurate model of the each other's emotions.  The
results show the opposite: peers in the audio-only condition built more accurate
models, probably because they concentrated more on the shared concept map. This
confirms other studies that revealed that viewing what one's partner sees
(shared graphical editor) is more important than seeing each
other~\citep{gaver1993one,anderson1997impact}. Therefore, what is meant by the
quality of interactions is $\mathcal{H}_{3}$ is more than the linguistic features
of dialogue but includes the way these interactions are articulated to the task.

@@

We found symmetry of the mutual model on 4 studies out of 5. In study 4, this
only applies to the control group (having read the same text before), which
corresponds to the situation of the 3 first studies.  Finding this symmetry per
se constitutes an interesting result as we don't know earlier studies who have
established this relationship. The problem is that this symmetry does not allow
use to discriminate  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The second
hypothesis, the skill of helping one's peer to model oneself, is questioned by
study 3 where C could be modelled  accurately by A and not accurately by B. The
same hypothesis is however supported by the results of study 2, and indirectly
by study 4.  This does not discard $\mathcal{H}_{1}$: even when we found
significant correlations, they were all below 0.50, not around 0.90.  Hence,
even of the quality of social interaction matters, there is obviously a large
part of individual variance within teams.
